Countering language attrition with PanLex and the Web of Data

### 2013-01-02 paper/2012/SWJ-PanLex/paper.tex
minor enhancements in linking section

    ?iso3 a plx:Iso639-3Code ;
          owl:sameAs ?lexvo3 .
  }
  With
    ?lang = uri(concat(plx:language, '/', ?iso3))
    ?iso3 = uri(concat(plx:iso639-3, '/', ?iso3))
    ?lexvo3 = uri(concat('http://lexvo.org/id/iso639-3/', ?iso3))
  From
    i1
\end{lstlisting}
\caption{An excerpt of a Sparqlify-ML view definition for PanLex's languages. This
example also demonstrates how ``is-a relations'' to schema.org and links to lexvo are established.}
\label{fig:ex:sparqlify-ml}
\end{figure}

\section{Linking}
\label{sec:linking}
The Sparqlify view in the previous section ~\ref{fig:ex:sparqlify-ml} already
established the interlinking of the PanLex languages with Lexvo.
In this section we outline the interlinking with DBpedia.

@@

%Since Lexvo covers all ISO 639-3 language codes, interlinking could be done with
%the Sparqlify-ML view in Figure~\ref{fig:ex:sparqlify-ml}.
For DBpedia we were interested in creating \emph{valid} and
thus \emph{dereferenceable} links.
Therefore, we iterated the \emph{titles} datasets\footnote{\url{http://wiki.dbpedia.org/Downloads38} - Note that more languages are available in the raw download section at \url{http://downloads.dbpedia.org/3.8/}},
which map (non-localized) DBpedia URIs to their page titles in the respective language.
For each language version we normalized the labels by applying
Unicode NFKD\footnote{\url{http://unicode.org/reports/tr15/}} normalization and
removal of punctuation characters.
% PanLex now uses NFC, not NFKD, for character normalization and does not remove punctuation.
Each DBpedia resource was then mapped to the PanLex expression that was equal to the resource's normalized label in the respective language.
Table ~\ref{fig:plx-dbp-link-counts} summarizes the number of links obtained.
\begin{table}
\centering
\begin{scriptsize}
\begin{tabular}{llll}
\toprule
Language & 639-1 & 639-3 & Links \\
\midrule
English    & en & eng & 1,415,241 \\
German     & de & deu &   224,146 \\
French     & fr & fra &   187,364 \\
Italian    & it & ita &   147,485 \\
Spanish    & sp & spa &   117,056 \\
Portuguese & pt & por &   112,266 \\
Polish     & pl & pol &   110,974 \\
Russian    & ru & rus &    68,040 \\
Czech      & cs & ces &    28,767 \\
Catalan    & ca & cat &    27,779 \\
Korean     & ko & kor &    24,912 \\
Turkish    & tr & tur &    22,258 \\
Bulgarian  & bg & bul &    19,431 \\
Hungarian  & hu & hun &    18,203 \\
Slovene    & sl & slv &    11,981 \\
Greek      & el & ell &     1,112 \\
\midrule
Total      &    &     & 2,537,015 \\
\bottomrule
\end{tabular}
\end{scriptsize}
\caption{Number of DBpedia links per language}
\label{fig:plx-dbp-link-counts}
\end{table}

>>>>

    ?iso3 a plx:Iso639-3Code ;
          owl:sameAs ?lexvo3 .
  }
  With
    ?lang = uri(concat(plx:language, '/', ?iso3))
    ?iso3 = uri(concat(plx:iso639-3, '/', ?iso3))
    ?lexvo3 = uri(concat('http://lexvo.org/id/iso639-3/', ?iso3))
  From
    i1
\end{lstlisting}
\caption{An excerpt of a Sparqlify-ML view definition for PanLex's languages. This
example also demonstrates how ``is-a relations'' to schema.org and links to Lexvo are established.}
\label{fig:ex:sparqlify-ml}
\end{figure}

\section{Linking}
\label{sec:linking}
The Sparqlify view in the previous section (\autoref{fig:ex:sparqlify-ml}) already
established the interlinking of the PanLex languages with Lexvo.
In this section we outline the interlinking with DBpedia.

@@

%Since Lexvo covers all ISO 639-3 language codes, interlinking could be done with
%the Sparqlify-ML view in Figure~\ref{fig:ex:sparqlify-ml}.
For DBpedia we were interested in creating \emph{valid} and
thus \emph{dereferenceable} links.
Therefore, we iterated the \emph{titles} datasets\footnote{\url{http://wiki.dbpedia.org/Downloads38} - Note that more languages are available in the raw download section at \url{http://downloads.dbpedia.org/3.8/}},
which map (non-localized) DBpedia URIs to their page titles in the respective language.
For each language version we normalized the labels by applying
Unicode NFKD\footnote{\url{http://unicode.org/reports/tr15/}} normalization and
removal of punctuation characters.
% PanLex now uses NFC, not NFKD, for character normalization and does not remove punctuation.
Each DBpedia resource was then mapped to the PanLex expression that was equal to the resource's normalized label in the respective language.
\autoref{fig:plx-dbp-link-counts} summarizes the number of links obtained.
\begin{table}
\centering
\begin{scriptsize}
\begin{tabular}{lllr}
\toprule
Language & 639-1 & 639-3 & Links \\
\midrule
English    & en & eng & 1,415,241 \\
German     & de & deu &   224,146 \\
French     & fr & fra &   187,364 \\
Italian    & it & ita &   147,485 \\
Spanish    & sp & spa &   117,056 \\
Portuguese & pt & por &   112,266 \\
Polish     & pl & pol &   110,974 \\
Russian    & ru & rus &    68,040 \\
Czech      & cs & ces &    28,767 \\
Catalan    & ca & cat &    27,779 \\
Korean     & ko & kor &    24,912 \\
Turkish    & tr & tur &    22,258 \\
Bulgarian  & bg & bul &    19,431 \\
Hungarian  & hu & hun &    18,203 \\
Slovene    & sl & slv &    11,981 \\
Greek      & el & ell &     1,112 \\
\midrule
Total      &    &     & 2,537,015 \\
\bottomrule
\end{tabular}
\end{scriptsize}
\caption{Number of DBpedia links per language}
\label{fig:plx-dbp-link-counts}
\end{table}
